%===============================================================================
% Member A — Implementation Guide
% Detailed sprint breakdown + how to approach each phase
%===============================================================================
\documentclass[12pt,a4paper]{report}
\input{_preamble.tex}

\pagestyle{fancy}
\fancyhf{}
\lhead{\textcolor{primary}{Member A Implementation Guide}}
\rfoot{Page \thepage}

\begin{document}

\begin{titlepage}
\centering
\vspace*{1cm}
{\Large\color{primary}\textbf{COMSATS University Islamabad (Lahore Campus)}}\\[0.3cm]
{\large Department of Computer Engineering}\\[1cm]
\rule{\textwidth}{2pt}\\[0.8cm]
{\Huge\bfseries\color{primary} Implementation Guide}\\[0.4cm]
{\Large Member A --- Platform, MQTT \& Deployment}\\[0.2cm]
{\large Detailed Sprint Breakdown \& Approach}\\[0.3cm]
\rule{\textwidth}{2pt}\\[1cm]
{\small Companion: \texttt{member\_a\_sprint\_plan.tex} $\cdot$ Master: \texttt{centralized\_sprint\_plan.tex}}\\
\vfill
{\large Spring--Fall 2026}
\end{titlepage}

\tableofcontents
\newpage

%===============================================================================
\chapter{Introduction}
%===============================================================================

\section{Purpose}

This guide tells you \textbf{how to implement} your track week by week --- not just \textit{what} to build. Use it alongside:

\begin{itemize}
    \item \texttt{.sprint/member\_a\_sprint\_plan.tex} --- task list
    \item \texttt{.sprint/centralized\_sprint\_plan.tex} --- joint dates \& milestones
    \item \texttt{.prerequisite/MEMBER\_A\_PREREQUISITES.md} --- skills before Sprint 1
    \item \texttt{finalization/dinosaur\_fyp\_architecture\_92471e72.plan.md} --- full architecture
\end{itemize}

\section{Your boundary}

You own everything from MQTT publish through to the dashboard. Member B stops at a Python handoff dict. \textbf{Never wait for Jetson until Sprint 4} --- the event simulator is your edge substitute in Sprints 1--3.

\section{General workflow every week}

\begin{enumerate}
    \item \textbf{Monday:} Pick Jira tickets for the week from active sprint
    \item \textbf{Daily:} Code on feature branch; push at least every 2 days
    \item \textbf{Wednesday:} 30-min sync with Member B (blockers only)
    \item \textbf{Friday:} Open/update PR; comment on Jira ticket
    \item \textbf{Sunday:} Update \texttt{docs/sprint-logs/} weekly entry
\end{enumerate}

\begin{tipbox}
Build \textbf{vertically thin slices} each sprint: make one event flow end-to-end before adding features. Example Sprint 2: one simulator alert $\rightarrow$ DB row $\rightarrow$ REST query --- before polish.
\end{tipbox}

%===============================================================================
\chapter{Sprint 1 (Weeks 1--4): Foundation \& Transport}
%===============================================================================

\section{Sprint goal}

Monorepo exists, contract drafted, Mosquitto + Redis + bridge + simulator running. Member B aligns handoff spec. \textbf{Milestone M0} at week 4.

\begin{approachbox}
\textbf{Order of work:}
\begin{enumerate}
    \item Dev environment + GitHub/Jira setup (Day 1--2)
    \item Repo scaffold + \texttt{contracts/events.py} (Week 1)
    \item Docker Compose with Mosquitto + Redis only (Week 2)
    \item Bridge + simulator (Week 3)
    \item Contract sign-off with B + integration session (Week 4)
\end{enumerate}
\textbf{Do not} start FastAPI services yet --- transport must work first.
\end{approachbox}

\subsection{Week 1--2: Scaffold \& contract}

\begin{stepbox}{Initialize monorepo}
Create top-level layout:
\begin{verbatim}
smart-surveillance/
|-- contracts/events.py
|-- transport/mqtt_redis_bridge/
|-- transport/event_simulator/
|-- edge/mqtt_publisher/        # you own; B never commits here
|-- services/alert_service/
|-- infra/docker-compose.yml
|-- frontend/src/
\end{verbatim}
Open Jira project SSP; create Sprint 1; tickets: SSP-1 scaffold, SSP-2 contract, SSP-3 compose.
\end{stepbox}

\begin{stepbox}{Draft frozen contract}
In \texttt{contracts/events.py} define Pydantic models:
\begin{itemize}
    \item \texttt{type}: alert | metrics | heartbeat
    \item \texttt{event}: helmet\_missing, vest\_missing, fire, smoke, spill, compliant
    \item Fields: timestamp, session\_id, worker\_id, zone, confidence, compliance\_score, detections[], image\_path
\end{itemize}
Share PDF/markdown mapping table with Member B: handoff dict field $\rightarrow$ contract field.
\end{stepbox}

\begin{stepbox}{Docker Compose skeleton}
Start minimal \texttt{infra/docker-compose.yml}:
\begin{itemize}
    \item Network \texttt{ppe-network}
    \item Services: \texttt{mosquitto}, \texttt{redis} (health checks)
    \item Volumes for Redis persistence
    \item \texttt{.env.example} with \texttt{MQTT\_HOST}, \texttt{REDIS\_URL}
\end{itemize}
Verify: \texttt{docker compose config} passes.
\end{stepbox}

\subsection{Week 3--4: Bridge \& simulator}

\begin{stepbox}{MQTT-Redis bridge}
Implement async Python service:
\begin{enumerate}
    \item Subscribe to \texttt{ppe/\#} on Mosquitto
    \item Parse JSON; validate with \texttt{contracts.events}
    \item Route: \texttt{ppe/alerts/*} $\rightarrow$ \texttt{XADD ppe:alerts}
    \item Add metadata: \texttt{received\_at}, \texttt{device\_id}
    \item On parse fail: \texttt{XADD ppe:dead\_letter}
    \item Exponential backoff reconnect
\end{enumerate}
Test: \texttt{mosquitto\_pub} manually $\rightarrow$ \texttt{XRANGE ppe:alerts}.
\end{stepbox}

\begin{stepbox}{Event simulator}
Publish realistic events at 45/hour:
\begin{itemize}
    \item All zones: welding\_bay, general, chemical\_storage
    \item All priorities; include compliant + violation mix
    \item Valid \texttt{image\_path} strings (no real files needed yet)
\end{itemize}
This replaces Jetson until Sprint 4.
\end{stepbox}

\begin{verifybox}
\begin{itemize}
    \item[$\square$] \texttt{docker compose up} starts mosquitto, redis, bridge, simulator
    \item[$\square$] \texttt{XRANGE ppe:alerts - + COUNT 10} shows events
    \item[$\square$] Bridge survives restart (Redis persistence)
    \item[$\square$] Member B signed handoff mapping doc (M0)
    \item[$\square$] First PR merged; Jira Sprint 1 tickets closed
\end{itemize}
\end{verifybox}

\begin{collabbox}
\textbf{Week 4 integration:} Walk Member B through contract fields. Receive sample handoff JSON from B; write a one-off script that converts it to contract payload (preview for Sprint 4 publisher).
\end{collabbox}

\begin{pitfallbox}
\begin{itemize}
    \item Starting microservices before bridge works --- debug transport in isolation first
    \item Changing contract after M0 without B's approval
    \item Committing \texttt{.env} secrets to GitHub
\end{itemize}
\end{pitfallbox}

%===============================================================================
\chapter{Sprint 2 (Weeks 5--8): Backend Core}
%===============================================================================

\section{Sprint goal}

PostgreSQL + MinIO live; Alert and Metrics services consume streams and expose REST. \textbf{Milestone M1} at week 8.

\begin{approachbox}
\textbf{Order of work:}
\begin{enumerate}
    \item Database schema + migrations (Week 5)
    \item Alert Service consumer only --- log to console (Week 5)
    \item Alert Service DB write + dedup (Week 6)
    \item Metrics Service + hypertable (Week 7)
    \item REST endpoints + integration test (Week 8)
\end{enumerate}
Keep simulator running entire sprint.
\end{approachbox}

\subsection{Week 5--6: Database \& Alert Service}

\begin{stepbox}{PostgreSQL schema}
Add to Compose: PostgreSQL + TimescaleDB image.
\begin{verbatim}
violations (id, timestamp, worker_id, zone,
  event_type, priority, confidence, image_path,
  agent_briefing, acknowledged)

compliance_metrics (time, zone, shift, score,
  worker_count, violation_count)  -- hypertable

zone_rules (zone, requires_helmet, requires_vest)
\end{verbatim}
Run migrations; seed \texttt{zone\_rules} for 3 zones.
\end{stepbox}

\begin{stepbox}{MinIO}
Add MinIO service; create \texttt{evidence/} bucket. Alert service will store/fetch paths --- upload test JPEG manually to verify presigned URLs.
\end{stepbox}

\begin{stepbox}{Alert Service (:8001)}
\begin{enumerate}
    \item Consumer group \texttt{alert-svc} on \texttt{ppe:alerts}
    \item Dedup: Redis key \texttt{dedup:\{worker\}:\{event\}:\{zone\}} TTL 60s
    \item INSERT into \texttt{violations}; \texttt{XACK} after success
    \item CRITICAL priority: skip rate limit
    \item Publish to Redis pub/sub channel \texttt{live:alerts} for Sprint 3 WebSocket
    \item \texttt{GET /api/alerts} with filters
\end{enumerate}
\end{stepbox}

\subsection{Week 7--8: Metrics Service}

\begin{stepbox}{Metrics Service (:8002)}
\begin{enumerate}
    \item Consumer \texttt{metrics-svc} on \texttt{ppe:metrics}
    \item Write to \texttt{compliance\_metrics} hypertable
    \item Rolling hourly/daily aggregates per zone
    \item Zone risk score: weighted violations $\times$ severity
    \item Endpoints: \texttt{/api/metrics/compliance}, \texttt{/trend}, \texttt{/zone-risk}
\end{enumerate}
\end{stepbox}

\begin{stepbox}{Integration test}
Pytest with docker-compose:
\begin{enumerate}
    \item Start stack + simulator
    \item Wait 30s; assert \texttt{violations} count $>$ 0
    \item \texttt{GET /api/alerts} returns 200
    \item \texttt{GET /api/metrics/compliance} returns data
\end{enumerate}
\end{stepbox}

\begin{verifybox}
\begin{itemize}
    \item[$\square$] M1: simulator $\rightarrow$ Redis $\rightarrow$ Alert $\rightarrow$ PostgreSQL
    \item[$\square$] REST returns filtered alerts
    \item[$\square$] Metrics hypertable has rows
    \item[$\square$] Ingested B's sample handoff JSON via test script (optional)
\end{itemize}
\end{verifybox}

\begin{collabbox}
\textbf{Week 8:} Integration session. Demo REST API to Member B. Collect any handoff JSON samples for Sprint 3 harness.
\end{collabbox}

%===============================================================================
\chapter{Sprint 3 (Weeks 9--12): Real-Time, Reports, Synthetic, Agent}
%===============================================================================

\section{Sprint goal}

WebSocket live; 90-day synthetic seed; PDF reports; agent skeleton with 3 tools. \textbf{M2} at week 12.

\begin{approachbox}
\textbf{Parallel tracks within sprint:}
\begin{itemize}
    \item \textbf{Track A:} WebSocket + Reporting (Week 9--10) --- demo value fast
    \item \textbf{Track B:} Synthetic data (Week 9--10) --- unblocks dashboard/agent testing
    \item \textbf{Track C:} Agent skeleton (Week 11--12) --- after DB has synthetic history
\end{itemize}
Review Member B's merged PRs in \texttt{edge/} but do not write CV code.
\end{approachbox}

\subsection{Week 9--10: WebSocket \& Reporting}

\begin{stepbox}{WebSocket Gateway (:8004)}
\begin{enumerate}
    \item Subscribe Redis \texttt{live:alerts} and metrics channels
    \item FastAPI \texttt{/ws/live}; connection manager (add/remove clients)
    \item Broadcast JSON on each alert
    \item Test with \texttt{websocat} or browser DevTools
\end{enumerate}
\end{stepbox}

\begin{stepbox}{Reporting Service (:8003)}
ReportLab PDF: header, date range, compliance summary, violation table, recommendations paragraph. \texttt{POST /api/reports/generate}. Scheduled cron optional stub.
\end{stepbox}

\begin{stepbox}{Synthetic data engine}
\texttt{synthetic/scenarios/90day\_baseline.yaml}: 15 workers, 3 zones, patterns (night drop, welding bay chronic violations). Run \texttt{python synthetic/generate.py} --- bulk INSERT + optional stream replay. Dashboard and agent now have history.
\end{stepbox}

\subsection{Week 11--12: Agent skeleton}

\begin{stepbox}{Agent Service (:8005)}
\begin{enumerate}
    \item LangGraph: planner $\rightarrow$ tool executor $\rightarrow$ responder
    \item Tools calling your DB: \texttt{query\_violations}, \texttt{get\_compliance\_trend}, \texttt{get\_zone\_risk\_score}
    \item \texttt{POST /api/agent/query} with OpenAI GPT-4o; Ollama fallback env flag
    \item OpenAPI on all 5 services
\end{enumerate}
\end{stepbox}

\begin{stepbox}{Handoff test harness}
Script: read B's handoff JSON file $\rightarrow$ your publisher function $\rightarrow$ MQTT $\rightarrow$ verify row in DB. No Jetson required.
\end{stepbox}

\begin{verifybox}
\begin{itemize}
    \item[$\square$] WebSocket client receives event within 2s of simulator publish
    \item[$\square$] PDF downloads with synthetic data charts
    \item[$\square$] Agent answers: ``How many violations in welding bay last 24h?''
    \item[$\square$] B's edge PRs merged; harness passes with B's sample dict
\end{itemize}
\end{verifybox}

%===============================================================================
\chapter{Sprint 4 (Weeks 13--16): Dashboard \& Jetson}
%===============================================================================

\section{Sprint goal}

Full React UI; MQTT publisher on Jetson wrapping B's code; agent RAG + reactive mode. \textbf{M3} at week 16.

\begin{approachbox}
\textbf{Order of work:}
\begin{enumerate}
    \item React shell + WebSocket + Dashboard page (Week 13) --- use synthetic data immediately
    \item Remaining pages (Week 14)
    \item \texttt{edge/mqtt\_publisher/} adapter for B's handoff (Week 15)
    \item Jetson flash + deploy + disable simulator (Week 15--16)
    \item Agent RAG + reactive CRITICAL briefings (Week 16)
\end{enumerate}
\end{approachbox}

\subsection{Week 13--14: React dashboard}

\begin{stepbox}{Frontend scaffold}
Vite + React + TypeScript + Zustand + Tailwind. WebSocket hook updates store. Pages:
\begin{itemize}
    \item Dashboard: compliance gauge, mini alert feed
    \item Alerts: full feed, filters, evidence modal
    \item Analytics: trend chart, zone heatmap, pie chart
    \item Agent: chat panel
    \item Reports: date picker $\rightarrow$ PDF
    \item Config: helmet/vest toggles per zone
\end{itemize}
\end{stepbox}

\subsection{Week 15--16: Jetson integration}

\begin{stepbox}{MQTT publisher on edge}
\texttt{edge/mqtt\_publisher/publisher.py}:
\begin{enumerate}
    \item Import B's \texttt{get\_compliance\_event(frame)} from compliance engine
    \item Map dict $\rightarrow$ \texttt{contracts.events.AlertPayload}
    \item Add timestamp, session\_id, priority, type
    \item paho-mqtt publish to broker (broker IP = your laptop/server on network)
\end{enumerate}
\end{stepbox}

\begin{stepbox}{Jetson deploy}
\begin{enumerate}
    \item Flash JetPack; SSH; install deps from B's \texttt{requirements.txt}
    \item Copy B's weights/ONNX; verify inference FPS
    \item Run publisher + inference loop as systemd service
    \item Point Jetson MQTT at Compose Mosquitto IP
    \item Verify end-to-end; stop simulator
\end{enumerate}
\end{stepbox}

\begin{stepbox}{Agent completion}
pgvector embeddings on violations; reactive trigger on CRITICAL stream event; write \texttt{agent\_briefing} to DB; push via WebSocket.
\end{stepbox}

\begin{verifybox}
\begin{itemize}
    \item[$\square$] M3: remove helmet on camera $\rightarrow$ gauge drops $\rightarrow$ alert on UI
    \item[$\square$] Agent briefing appears on CRITICAL alert
    \item[$\square$] All 6 pages functional
\end{itemize}
\end{verifybox}

\begin{collabbox}
\textbf{Weeks 15--16:} B available for threshold tuning questions. You operate Jetson during integration session.
\end{collabbox}

%===============================================================================
\chapter{Sprint 5 (Weeks 17--20): DevOps \& Demo}
%===============================================================================

\section{Sprint goal}

One-command stack; CI/CD; Grafana; 12-minute demo rehearsed. \textbf{M4} at week 20.

\begin{approachbox}
Do DevOps \textbf{after} features work --- don't block Sprint 4 UI on Grafana. Week 17--18: Compose hardening + CI. Week 19--20: demo script + report.
\end{approachbox}

\begin{stepbox}{Full docker-compose}
All services with \texttt{depends\_on} health conditions. Nginx serves React build. Document \texttt{README.md} quick start.
\end{stepbox}

\begin{stepbox}{GitHub Actions}
Lint (ruff, eslint) $\rightarrow$ pytest $\rightarrow$ compose integration job $\rightarrow$ build images.
\end{stepbox}

\begin{stepbox}{Observability}
Prometheus scrape all \texttt{/metrics}. Grafana dashboards: System Health + Safety KPIs. Show both during demo.
\end{stepbox}

\begin{stepbox}{12-minute demo script}
\begin{enumerate}
    \item \texttt{docker compose up} (1 min)
    \item Grafana health (1 min)
    \item Live violation + agent briefing (2 min)
    \item Agent chat question (2 min)
    \item Analytics 90-day view (2 min)
    \item PDF report (1 min)
    \item Zone config change (1 min)
    \item Q\&A architecture (2 min)
\end{enumerate}
Rehearse twice; record screen backup.
\end{stepbox}

\begin{verifybox}
\begin{itemize}
    \item[$\square$] M4: demo run without manual fixes
    \item[$\square$] CI green on \texttt{main}
    \item[$\square$] Platform report chapters drafted
    \item[$\square$] Tag \texttt{v1.0-demo}
\end{itemize}
\end{verifybox}

%===============================================================================
\chapter{Appendix A: Service Ports}
%===============================================================================

\begin{center}
\begin{tabular}{cl}
\toprule
\textbf{Port} & \textbf{Service} \\
\midrule
1883 & Mosquitto \\
6379 & Redis \\
5432 & PostgreSQL \\
9000 & MinIO \\
8001--8005 & FastAPI services \\
3000 & React (Nginx) \\
9090 / 3001 & Prometheus / Grafana \\
\bottomrule
\end{tabular}
\end{center}

%===============================================================================
\chapter{Appendix B: Troubleshooting}
%===============================================================================

\begin{itemize}
    \item \textbf{No stream messages:} Check bridge logs; \texttt{XRANGE}; Mosquitto ACL
    \item \textbf{Consumer lag:} \texttt{XPENDING ppe:alerts alert-svc}; claim stale messages
    \item \textbf{WebSocket silent:} Verify Redis pub/sub from Alert Service
    \item \textbf{Jetson cannot reach broker:} Same LAN IP; firewall port 1883
    \item \textbf{Agent empty answers:} Check synthetic seed; OpenAI key; tool SQL
\end{itemize}

\end{document}
